Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-08 arxiv/intro.tex
abstract

\begin{abstract}
	Multi-task learning is a powerful approach in many applications such as image and text classification. Yet, there is little rigorous understanding of when multi-task learning outperforms single-task learning. In this work, we provide a rigorous study to answer the question in the high-dimensional linear regression setting. We show that a bias-variance tradeoff of multi-task learning determines the effect of information transfer, and develop new concentration bounds to analyze the tradeoff. Our key observation is that three properties of task data, namely \textit{task similarity}, \textit{sample ratio}, and \textit{covariate shift} can affect transfer in the high-dimensional linear regression setting. We relate each property to the bias and variance of multi-task learning and explain three negative effects with decreased task similarity,	increased sample ratio, and covariate shift under increased sample ratio. We validate the three effects on text classification tasks. Inspired by our theory, we show two practical connections of interest.
	First, single-task results can help to understand when multi-task learning gives gains. Second, incrementally adding training data can mitigate negative transfer and improve multi-task training efficiency.
\end{abstract}

@@

%Multi-task learning is an inductive learning mechanism to improve generalization performance using related task data.
%Many state-of-the-art results in computer vision and natural language processing are obtained using multi-task learning.
Multi-task learning is a powerful approach to improve performance for many tasks in computer vision, natural language processing, and other areas \cite{C97,ZY17,R17}.
%In multi-task learning, having related task data is fundamental to its performance.
%Multi-task learning is particularly powerful when there is limited labeled data for a task to be solved, meanwhile more labeled data from different but related tasks is available.
%By combining multiple information sources, it is possible to share all the information in the same model.
In many settings, multiple source tasks are available to help with predicting a particular target task.
\todo{clarify setting is different from traditional MTL}
%For example, many applications in , and many other areas have been achieved by learning from multiple tasks together.
The performance of multi-task learning depends on the relationship between the source and target tasks \cite{C97}.
%	We define that multi-task learning provides \textit{positive transfer} if it outperforms single-task learning, or \textit{negative transfer} otherwise.
When the sources are relatively different from the target, multi-task learning (MTL) has often been observed to perform worse than single-task learning (STL) \cite{AP16,BS17}, which is referred to as \textit{negative transfer} \cite{PY09}.
While many empirical approaches have been proposed to mitigate negative transfer \cite{ZY17}, a precise understanding of when negative transfer occurs remains elusive in the literature \cite{R17}.
%This phenomenon, known as \textit{negative transfer}, is fundamental to the understanding of multi-task learning.

@@

%Let $n_i = \rho_i \cdot p$ denote the data size and $X_i\in\real^{n_i\times p}$ denote the features of task $i$.
%The labels of task $i$ are given by $Y_i = X_i\beta_i + \varepsilon_i$, where $\beta_i\in\real^p$ denotes task $i$'s ground truth parameters and $\varepsilon_i$ denotes i.i.d. random noise with mean zero and variance $\sigma^2$.
We consider a multi-task estimator obtained using a shared linear layer for all tasks and a separate output layer for each task \cite{WZR20}.
This two-layer model is inspired by a commonly used idea of hard parameter sharing in multi-task learning \cite{R17,MTDNN19}.
We consider the bias and variance of the multi-task estimator for predicting a target task and compare its performance to single-task learning.
%This corresponds to minimizing $ \sum_{i=1}^t \norm{X_i B W_i - Y_i}^2$.
%Let $\hat{\beta}_t^{\MTL}$ denote the optimal multi-task estimator for the target task, which is defined precisely in Section \ref{sec_prelim}.
%We revisit the bias-variance tradeoff of $\hat{\beta}_t^{\MTL}$.

>>>>

\begin{abstract}
	Hard parameter sharing is a widely used approach to learn from multiple tasks in many applications, such as text classification.
	The idea is that all tasks share the same feature layer, while each task uses a specific output layer for prediction.
	Intuitively, the performance of hard parameter sharing depends on dataset properties such as their sample sizes.
	Yet, rigorously formulating such intuition can be technically challenging.
	This paper analyzes the generalization performance of hard parameter sharing in high-dimensional linear regression, where the sample size and feature dimension become increasingly large in a fixed ratio.
	We present tight generalization bounds
	%that scale with datasets' sample sizes and the spectrum of their feature covariates
	for two canonical cases:
	(i) multiple tasks with the same feature covariates;
	(ii) two tasks with arbitrarily different sample sizes and covariance matrices.
	We also demonstrate that these estimates for the empirical loss of hard parameter sharing are incredibly accurate in modest dimensions.
	Finally, we apply our results to explain several intriguing phenomena; for example, increasing one task's sample size helps another task initially by reducing variance, but hurts eventually due to increasing bias.
	This suggests progressively adding data as a more efficient procedure for optimizing hard parameter sharing, and we validate its efficiency in a text classification task.
%	Multi-task learning is a powerful approach in many applications such as image and text classification. Yet, there is little rigorous understanding of when multi-task learning outperforms single-task learning. In this work, we provide a rigorous study to answer the question in the high-dimensional linear regression setting. We show that a bias-variance tradeoff of multi-task learning determines the effect of information transfer, and develop new concentration bounds to analyze the tradeoff. Our key observation is that three properties of task data, namely \textit{task similarity}, \textit{sample ratio}, and \textit{covariate shift} can affect transfer in the high-dimensional linear regression setting. We relate each property to the bias and variance of multi-task learning and explain three negative effects with decreased task similarity,	increased sample ratio, and covariate shift under increased sample ratio. We validate the three effects on text classification tasks. Inspired by our theory, we show two practical connections of interest.
%	First, single-task results can help to understand when multi-task learning gives gains. Second, incrementally adding training data can mitigate negative transfer and improve multi-task training efficiency.
\end{abstract}

@@

\iffalse
%Multi-task learning is an inductive learning mechanism to improve generalization performance using related task data.
%Many state-of-the-art results in computer vision and natural language processing are obtained using multi-task learning.
Multi-task learning is a powerful approach to improve performance for many tasks in computer vision, natural language processing, and other areas \cite{C97,ZY17,R17}.
%In multi-task learning, having related task data is fundamental to its performance.
%Multi-task learning is particularly powerful when there is limited labeled data for a task to be solved, meanwhile more labeled data from different but related tasks is available.
%By combining multiple information sources, it is possible to share all the information in the same model.
In many settings, multiple source tasks are available to help with predicting a particular target task.
\todo{clarify setting is different from traditional MTL}
%For example, many applications in , and many other areas have been achieved by learning from multiple tasks together.
The performance of multi-task learning depends on the relationship between the source and target tasks \cite{C97}.
%	We define that multi-task learning provides \textit{positive transfer} if it outperforms single-task learning, or \textit{negative transfer} otherwise.
When the sources are relatively different from the target, multi-task learning (MTL) has often been observed to perform worse than single-task learning (STL) \cite{AP16,BS17}, which is referred to as \textit{negative transfer} \cite{PY09}.
While many empirical approaches have been proposed to mitigate negative transfer \cite{ZY17}, a precise understanding of when negative transfer occurs remains elusive in the literature \cite{R17}.
%This phenomenon, known as \textit{negative transfer}, is fundamental to the understanding of multi-task learning.

@@

%Let $n_i = \rho_i \cdot p$ denote the data size and $X_i\in\real^{n_i\times p}$ denote the features of task $i$.
%The labels of task $i$ are given by $Y_i = X_i\beta_i + \varepsilon_i$, where $\beta_i\in\real^p$ denotes task $i$'s ground truth parameters and $\varepsilon_i$ denotes i.i.d. random noise with mean zero and variance $\sigma^2$.
We consider a multi-task estimator obtained using a shared linear layer for all tasks and a separate output layer for each task \cite{WZR20}.
This two-layer model is inspired by a commonly used idea of hard parameter sharing in multi-task learning \cite{R17,MTDNN19}.
We consider the bias and variance of the multi-task estimator for predicting a target task and compare its performance to single-task learning.
%This corresponds to minimizing $ \sum_{i=1}^t \norm{X_i B W_i - Y_i}^2$.
%Let $\hat{\beta}_t^{\MTL}$ denote the optimal multi-task estimator for the target task, which is defined precisely in Section \ref{sec_prelim}.
%We revisit the bias-variance tradeoff of $\hat{\beta}_t^{\MTL}$.
\fi
